---
title: "Week 6 2021 Recap"
author: "Sam Kirschner"
date: "2021-10-21T13:09:13-06:00"
output: html_document
---
 
## Week 6 Recap: SWAB roll to 5 game winning streak
 
This week didn't see much in the way of any unlucky losses, with the highest probability of win for a loser being Elliot's Warriors at 0.53. The Slugs and SWAB put in a dominating performance overall, while Titan's, Kraken, and the Sunday Winners all topped 90 points to bring home the win.

The one bit of luck exhibited this week was the Supermarket Studs facing the sputtering Bluecats. The Studs dropped a measly 80 points (0.42 win probability), but it was enough to pull out the win. The Bluecats with a wounded Russ Wilson looked like a shell of the team that started off the season with a 129 point win. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ffscrapr)
library(dplyr)
library(ggplot2)
library(gt)
library(sqldf)
library(gtExtras)
library(tidymodels)
library(tidyr)
library(DBI)
library(ggimage)



#### INPUT WHAT WEEK IT IS HERE ####
week_num<-6

swid<-"{FB82F2B3-89EB-4E1A-9677-B1333054B883}"

ESPN_S2<-"AECkrDCmS9dG4sssE55%2BNQRasW7jwjJVFyEUw6mR2wLFylIkvGy89sVJxc8GMvORMKJaBksVZfGnYB4BXKiM80pkw352zpsk%2BEbMg2ILadghIeMp2iz1Etpke3FMM2pyVdNg23edB2DoYIHMoUT72KS3Ho9bKNLPkkJ761JMh4o4ikTDTeL1uBJYNiahAMg1uvT4mHvszAP1UoDHQGLUPuOLchpl8Os8BsibtvBOTn778ksyx6gU513HRQGrpge7iQ9G04xOeWwYAjXfwcxa%2FCQr3Gx9kECPmidN9%2F7PrHImFQ%3D%3D"

espn_conn_2021 <- espn_connect(
  season = 2021,
  league_id = 214888,
  espn_s2 = ESPN_S2,
  swid = swid
)

kfl_schedule_2021<-ff_schedule(espn_conn_2021)


kfl_franchises_2021<-ff_franchises(espn_conn_2021)
kfl_franchises_2021$logo[10]<-"https://image.shutterstock.com/image-vector/lightning-warrior-logo-template-design-260nw-1019248687.jpg"
kfl_franchises_2021$logo[3]<-"https://www.pngkit.com/png/detail/269-2691865_winners-logo-winners.png"

kfl_db <-"kfl_ff_data" 
host_db <- "localhost"
db_port <- "5432"
db_user <- "postgres"  
db_password <- "Cincinnati1"
kfl_con <- dbConnect(RPostgres::Postgres(), dbname = kfl_db, host=host_db, port=db_port, user=db_user, password=db_password)   

kfl_schedule<-dbReadTable(kfl_con,"kfl_schedule")
kfl_franchises<-dbReadTable(kfl_con,"kfl_franchises")

```



```{r wins over expected model run, echo=FALSE}



score_distribution_data<-sqldf(
"select a.franchise_id,b.user_name,a.Years,a.week,franchise_score as score_for, opponent_score as score_against
 from kfl_schedule a
 left outer join kfl_franchises b 
 on a.franchise_id = b.franchise_id and a.Years = b.Years
 where week < 15"
)

score_distribution_data<-score_distribution_data%>%unite(YR_WK, c("Years","week"),remove = FALSE)

score_distribution_data<-score_distribution_data%>%group_by(YR_WK)%>%mutate(avg_weekly_score=mean(score_for),win = case_when(score_for > score_against ~ 1,score_for < score_against ~ 0))

score_distribution_data$win<-as.factor(score_distribution_data$win)

fitted_logistic_model<- logistic_reg() %>%
        # Set the engine
        set_engine("glm") %>%
        # Set the mode
        set_mode("classification") %>%
        # Fit the model
        fit(win~score_for, data = score_distribution_data)

```

```{r wins over expected model, echo=FALSE,warning=FALSE}
## 2021 win probs

kfl_2021_win_prob_data<-kfl_schedule_2021%>%rename( score_for = franchise_score)%>%filter(week<=week_num)

pred_prob_2021 <- predict(fitted_logistic_model,
                          new_data = kfl_2021_win_prob_data,
                          type = "prob")

colnames(pred_prob_2021)<-c("loss_prob", "win_prob")

kfl_2021_win_prob_data<-cbind(kfl_2021_win_prob_data,pred_prob_2021)

kfl_2021_win_prob_data<-kfl_2021_win_prob_data%>%mutate(win = case_when(score_for > opponent_score ~ 1,score_for < opponent_score ~ 0))

kfl_2021_win_prob_data[37,9]<-1
kfl_2021_win_prob_data[40,9]<-0
kfl_2021_win_prob_data[39,9]<-1
kfl_2021_win_prob_data[44,9]<-0


kfl_2021_weekly_data<-sqldf("SELECT A.*,B.franchise_name,B.franchise_abbrev,B.logo,C.franchise_name as opponent_franchise,C.franchise_abbrev as opponent_abbrev,C.logo as opponent_logo
      FROM kfl_2021_win_prob_data A
      LEFT OUTER JOIN kfl_franchises_2021 B
      on a.franchise_id = b.franchise_id
      LEFT OUTER JOIN kfl_franchises_2021 C
      on a.opponent_id = c.franchise_id")
```

```{r wk4 review, echo=FALSE,warning=FALSE,,message=FALSE}

wk4_matchups<-kfl_2021_weekly_data%>%filter(week==week_num &win==1)


############################## UPDATE THE WEEK #####################################################
wk4_matchups<-sqldf("select A.*,B.win_prob as opp_win_prob
                    FROM wk4_matchups A
                    LEFT OUTER JOIN kfl_2021_weekly_data B
                    ON A.opponent_id=B.franchise_id
                    WHERE B.week = 6")

wk4_matchups<-wk4_matchups%>%select(franchise_name,logo,win_prob,score_for,opponent_score,opp_win_prob,opponent_logo,opponent_franchise)


########################## UPDATE THE TITLE OF THE TAB HEADER #######################################
wk4_matchups%>%mutate(win_prob=round(win_prob,2),opp_win_prob=round(opp_win_prob,2))%>%gt()%>%gt_img_rows(logo)%>%gt_img_rows(opponent_logo)%>%
  tab_header(title ="Week 6 matchups")%>%
  cols_label(franchise_name="Winning Team",logo="",win_prob= "Win Probability",score_for="Score",opponent_score="Score",opp_win_prob="Win Probability",opponent_logo="",opponent_franchise="Losing Team")%>%
  tab_style(
    style = list(
      cell_borders(
        sides = "left",
        color = "black",
        weight = px(3)
      )
    ),
    locations = list(
      cells_body(
        columns = vars(opponent_score)
      )
    )
  ) %>%
  # We use tab_style() to change style of cells
  # cell_borders() provides the formatting
  # locations tells it where
  # Add black borders to the bottom of all the column labels
  tab_style(
    style = list(
      cell_borders(
        sides = "bottom",
        color = "black",
        weight = px(3)
      )
    ),
    locations = list(
      cells_column_labels(
        columns = gt::everything()
      )
    )
  )%>%
   data_color(
    columns = vars(score_for),
    colors = scales::col_numeric(
      palette = c("#c9ecb4"),
      domain = NULL
    ))%>%
   data_color(
    columns = vars(win_prob),
    colors = scales::col_numeric(
      palette = c("#ffffff", "#f2fbd2", "#c9ecb4", "#93d3ab", "#35b0ab"),
      domain = NULL
    ))%>%
   data_color(
    columns = vars(opponent_score),
    colors = scales::col_numeric(
      palette = c("#EF9A9AFF"),
      domain = NULL
    ))%>%
   data_color(
    columns = vars(opp_win_prob),
    colors = scales::col_numeric(
      # Using a function from paletteer to generate a vector of colors
      # Note that you need to wrap paletteer_d outputs in as.character()
      palette = as.character(paletteer::paletteer_d("ggsci::red_material", n = 5)),
      # Domain is from scales::col_numeric
      # Domain = The possible values that can be mapped
      # We don't HAVE to set a range for this since
      # we're using it inside data_color()
      domain = NULL
    ))

```

## Expected Wins power rankings through week 6

The Sunday Winners continue to show dominance with 4.2 expected wins on the season, but the Kraken, SWAB, and Slugs are hot on their heels now with roughly 4 expected wins a piece. In the next tier, the Supermarket Studs are moving up the ranks with the Tenderloins at 3.7 expected wins a piece. 

While the unlucky ducks should have about 2.4 more wins than they currently have given the quality of their team, they still sit last in their division and will find time is slipping to dig themselves out of this hole.

The Glum division is all in the bottom half of the power rankings with the Titans and Onionheads tied at 3.4 expected wins a piece and the Titan's have a 1 game lead on the Onionheads, but it is surely anyone's division.

Finally rounding out the basement is the GMOs, Warriors, and Bluecats. All three teams will need to get it going if they want to reverse their fortunes this season.


```{r expected_wins, echo=FALSE,warning=FALSE,fig.width = 10,fig.height=8,message=FALSE}
kfl_wins_over_expected<-kfl_2021_weekly_data%>%select(win_prob,win,franchise_id,franchise_name,logo,franchise_abbrev)%>%group_by(franchise_id,franchise_name,franchise_abbrev,logo)%>%summarise(expected_wins = sum(win_prob),wins = sum(win))%>%mutate(wins_over_expected = wins-expected_wins)%>%arrange(expected_wins)

week_4_table<-kfl_wins_over_expected%>%select(franchise_id,franchise_name,franchise_abbrev,logo,wins,expected_wins,wins_over_expected)%>%mutate(expected_wins = round(expected_wins,1), wins_over_expected = round(wins_over_expected,1))%>%arrange(desc(expected_wins))

week_4_table<-as.data.frame(week_4_table[,c(2:7)])

asp_ratio <- 1.618 


################### UPDATE PLOT TITLE ##############################
week4_exp_wins<-ggplot(kfl_wins_over_expected,aes(x=reorder(franchise_abbrev,expected_wins),y=expected_wins))+geom_bar(stat="identity")+geom_image(aes(x=franchise_abbrev,y=expected_wins,image=logo), size = 0.05, by = "width", asp = asp_ratio
)+coord_flip()+ggtitle("Expected Wins through Week 6") + xlab("franchise") + ylab("expected wins")+theme(axis.title.x = element_text(size=14),axis.title.y=element_text(size=14),axis.text = element_text(size = 12),plot.title = element_text(size = 18))

week4_exp_wins

################### UPDATE TABLE TITLE ##############################
week_4_table%>%select(franchise_name,logo,expected_wins,wins_over_expected)%>%gt()%>%gt_img_rows(logo) %>%tab_header(title ="Expected Wins through Week 6")%>%
  cols_label(franchise_name="Team Name",logo="",expected_wins="Expected Wins",wins_over_expected="Wins Over Expected")


#theme(aspect.ratio = 1/asp_ratio)

 #include aspect ratio in ggsave
#ggsave(
#  "week6_exp_wins.png", week4_exp_wins, 
#  # make the width equivalent to the aspect.ratio
#  height = 20, width = 20 * asp_ratio, dpi = "retina")






```
## Standings through Week 6

The Warriors and Bluecats will both look to break a 4 game losing streak this upcoming week or their seasons will be essentially over. It is truly do or die at this point in the season. 

Meanwhile, the SWAB look to continue their dominance and add another notch to their 5 game winning streak. Scott's TITANS will look to create some separation from the Onionheads and keep their 4 game streak alive.

The division to watch over the next couple weeks will be the Indiana division. The Kraken and Studs both sit at 4-2 with the tenderloins hot on their heels at 3-3. Will we see some separation and get an idea of who is in the hunt for the division and who will be forced to chase the wild card? 



```{r divison records, echo=FALSE,warning=FALSE,fig.width = 10,fig.height=8,message=FALSE}

kfl_standings_wk4<-kfl_2021_weekly_data%>%filter(week<=week_num)

franchise_id<-1:12
division_name<-c("Ohio","Glum","Ohio","Indiana","Glum","Ohio","Ohio","Indiana","Indiana","Glum","Glum","Indiana")
kfl_division<-as.data.frame(cbind(franchise_id,division_name))
kfl_division$franchise_id<-as.numeric(kfl_division$franchise_id)


kfl_division_data<- sqldf("Select A.*,B.division_name
      from kfl_standings_wk4 A
      left outer join kfl_division B
      on A.Franchise_ID = B.Franchise_ID
      ")

team_df<-kfl_division_data%>%filter(week == 1)%>%select(franchise_id,franchise_name,logo,division_name)

joined_df<-kfl_division_data%>%select(franchise_id,franchise_name,logo,division_name,win)%>%group_by(franchise_name,franchise_id)%>%summarise(
         Wins = length(win[win==1]),
         Losses = length(win[win==0]),
         outcomes = list(win), .groups = "drop") %>% 
     left_join(team_df, by = c("franchise_id" = "franchise_id","franchise_name"="franchise_name")) %>% 
     select(logo,franchise_name, division_name, Wins:outcomes)
 

###################### NEED TO UPDATE THE TITLE HEADER FOR WEEK  ##########################
standings_df<-joined_df%>%select(franchise_name,logo,division_name,Wins,Losses,outcomes)%>%gt(groupname_col="division_name")%>%gt_plt_winloss(outcomes,max_wins = 14)%>%gt_img_rows(logo)%>%tab_header(title ="Standings Through Week 6")%>%cols_label(franchise_name="Team",logo="",outcomes="Win-Loss Streak")

standings_df

```

