---
title: "Week 4 2022 Recap"
author: "Samuel Kirschner"
date: "2022-10-03T13:09:13-06:00"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ffscrapr)
library(dplyr)
library(ggplot2)
library(gt)
library(sqldf)
library(gtExtras)
library(tidymodels)
library(tidyr)
library(DBI)
library(ggimage)
#remotes::install_github("jthomasmock/espnscrapeR")
library(espnscrapeR)



#### INPUT WHAT WEEK IT IS HERE ####
week_num<-4

swid<-"{FB82F2B3-89EB-4E1A-9677-B1333054B883}"

ESPN_S2<-"AECkrDCmS9dG4sssE55%2BNQRasW7jwjJVFyEUw6mR2wLFylIkvGy89sVJxc8GMvORMKJaBksVZfGnYB4BXKiM80pkw352zpsk%2BEbMg2ILadghIeMp2iz1Etpke3FMM2pyVdNg23edB2DoYIHMoUT72KS3Ho9bKNLPkkJ761JMh4o4ikTDTeL1uBJYNiahAMg1uvT4mHvszAP1UoDHQGLUPuOLchpl8Os8BsibtvBOTn778ksyx6gU513HRQGrpge7iQ9G04xOeWwYAjXfwcxa%2FCQr3Gx9kECPmidN9%2F7PrHImFQ%3D%3D"

espn_conn_2022 <- espn_connect(
  season = 2022,
  league_id = 214888,
  espn_s2 = ESPN_S2,
  swid = swid
)

kfl_schedule_2021<-ff_schedule(espn_conn_2022)
kfl_pscores<-ff_playerscores(espn_conn_2022)
starters<-ff_starters(espn_conn_2022)


kfl_franchises_2021<-ff_franchises(espn_conn_2022)
#kfl_franchises_2021$logo[10]<-"https://image.shutterstock.com/image-vector/lightning-warrior-logo-template-design-260nw-1019248687.jpg"
kfl_franchises_2021$logo[3]<-"https://www.pngkit.com/png/detail/269-2691865_winners-logo-winners.png"

kfl_db <-"kfl_ff_data" 
host_db <- "localhost"
db_port <- "5432"
db_user <- "postgres"  
db_password <- "Cincinnati1"
kfl_con <- dbConnect(RPostgres::Postgres(), dbname = kfl_db, host=host_db, port=db_port, user=db_user, password=db_password)   

kfl_schedule<-dbReadTable(kfl_con,"kfl_schedule")
kfl_franchises<-dbReadTable(kfl_con,"kfl_franchise")

```



```{r wins over expected model run, echo=FALSE}



score_distribution_data<-sqldf(
"select a.franchise_id,b.user_name,a.Years,a.week,franchise_score as score_for, opponent_score as score_against
 from kfl_schedule a
 left outer join kfl_franchises b 
 on a.franchise_id = b.franchise_id and a.Years = b.Years
 where week < 15"
)

score_distribution_data<-score_distribution_data%>%unite(YR_WK, c("Years","week"),remove = FALSE)

score_distribution_data<-score_distribution_data%>%group_by(YR_WK)%>%mutate(avg_weekly_score=mean(score_for),win = case_when(score_for > score_against ~ 1,score_for < score_against ~ 0))

score_distribution_data$win<-as.factor(score_distribution_data$win)

fitted_logistic_model<- logistic_reg() %>%
        # Set the engine
        set_engine("glm") %>%
        # Set the mode
        set_mode("classification") %>%
        # Fit the model
        fit(win~score_for, data = score_distribution_data)

```

```{r wins over expected model, echo=FALSE,warning=FALSE}
## 2021 win probs

kfl_2021_win_prob_data<-kfl_schedule_2021%>%rename( score_for = franchise_score)%>%filter(week<=week_num)

pred_prob_2021 <- predict(fitted_logistic_model,
                          new_data = kfl_2021_win_prob_data,
                          type = "prob")

colnames(pred_prob_2021)<-c("loss_prob", "win_prob")

kfl_2021_win_prob_data<-cbind(kfl_2021_win_prob_data,pred_prob_2021)

kfl_2021_win_prob_data<-kfl_2021_win_prob_data%>%mutate(win = case_when(score_for > opponent_score ~ 1,score_for < opponent_score ~ 0))

#kfl_2021_win_prob_data[37,9]<-1
#kfl_2021_win_prob_data[40,9]<-0
#kfl_2021_win_prob_data[39,9]<-1
#kfl_2021_win_prob_data[44,9]<-0
kfl_2021_win_prob_data[41,9]<-0
kfl_2021_win_prob_data[45,9]<-1



kfl_2021_weekly_data<-sqldf("SELECT A.*,B.franchise_name,B.franchise_abbrev,B.logo,C.franchise_name as opponent_franchise,C.franchise_abbrev as opponent_abbrev,C.logo as opponent_logo
      FROM kfl_2021_win_prob_data A
      LEFT OUTER JOIN kfl_franchises_2021 B
      on a.franchise_id = b.franchise_id
      LEFT OUTER JOIN kfl_franchises_2021 C
      on a.opponent_id = c.franchise_id")
```
## Week 4 Recap

Almost half the league put up 100 points or more this week with some big performances throughout the KFL. The Bluecats finally busted through with a stud performance from Austin Ekeler and MVP Tom looking like his old self as they improved to 3-1 over the winless GMO's. The Onionheads roll on with a 103 point performance to go to 3-1 as the Monarchs have a middling performance and drop to 1-3. As for our prior unbeaten teams, The "UN" Lucky Ducks put up a big day with 110 points and are now the only unbeaten team in the KFL looking like sure favorites to make the playoffs. Elliot's Last Good Tight Ends live and die with the Ravens offense and this week thy died with a pitiful 45 point performance that saw them fall to the winless Supermarket Duds. Finally, both the Slugs and Kraken continue to chug forward below the radar with solid performances to improve to 3-1 respectively

```{r wk4 review, echo=FALSE,warning=FALSE,,message=FALSE}

wk4_matchups<-kfl_2021_weekly_data%>%filter(week==week_num &win==1)


############################## UPDATE THE WEEK #####################################################
wk4_matchups<-sqldf("select A.*,B.win_prob as opp_win_prob
                    FROM wk4_matchups A
                    LEFT OUTER JOIN kfl_2021_weekly_data B
                    ON A.opponent_id=B.franchise_id
                    WHERE B.week = 4")

wk4_matchups<-wk4_matchups%>%select(franchise_name,logo,win_prob,score_for,opponent_score,opp_win_prob,opponent_logo,opponent_franchise)


########################## UPDATE THE TITLE OF THE TAB HEADER #######################################
wk4_matchups%>%mutate(win_prob=round(win_prob,2),opp_win_prob=round(opp_win_prob,2))%>%gt()%>%gt_img_rows(logo)%>%gt_img_rows(opponent_logo)%>%
  tab_header(title ="Week 4 2022 matchups")%>%
  cols_label(franchise_name="Winning Team",logo="",win_prob= "Win Probability",score_for="Score",opponent_score="Score",opp_win_prob="Win Probability",opponent_logo="",opponent_franchise="Losing Team")%>%
  tab_style(
    style = list(
      cell_borders(
        sides = "left",
        color = "black",
        weight = px(3)
      )
    ),
    locations = list(
      cells_body(
        columns = vars(opponent_score)
      )
    )
  ) %>%
  # We use tab_style() to change style of cells
  # cell_borders() provides the formatting
  # locations tells it where
  # Add black borders to the bottom of all the column labels
  tab_style(
    style = list(
      cell_borders(
        sides = "bottom",
        color = "black",
        weight = px(3)
      )
    ),
    locations = list(
      cells_column_labels(
        columns = gt::everything()
      )
    )
  )%>%
   data_color(
    columns = vars(score_for),
    colors = scales::col_numeric(
      palette = c("#c9ecb4"),
      domain = NULL
    ))%>%
   data_color(
    columns = vars(win_prob),
    colors = scales::col_numeric(
      palette = c("#ffffff", "#f2fbd2", "#c9ecb4", "#93d3ab", "#35b0ab"),
      domain = NULL
    ))%>%
   data_color(
    columns = vars(opponent_score),
    colors = scales::col_numeric(
      palette = c("#EF9A9AFF"),
      domain = NULL
    ))%>%
   data_color(
    columns = vars(opp_win_prob),
    colors = scales::col_numeric(
      # Using a function from paletteer to generate a vector of colors
      # Note that you need to wrap paletteer_d outputs in as.character()
      palette = as.character(paletteer::paletteer_d("ggsci::red_material", n = 5)),
      # Domain is from scales::col_numeric
      # Domain = The possible values that can be mapped
      # We don't HAVE to set a range for this since
      # we're using it inside data_color()
      domain = NULL
    ))

```

## Expected Wins power rankings through week 4 2022

With the LGTE's terrible performance this week the Onionhead's sneak to the top of our expected wins leaderboard. The 3 through 6 slots are all the Ohio division with the Monarchs a full win under expected and the Bluecat's continuing to ride the wave with 1.25 wins over expected.That isn't the luckiest team in the league though which would go to the misnamed Unlucky ducks who are 1.4 wins over expected and sit in the bottom half of the power ranks. Finally, the Tenderloins are the only franchise yet to accumulate a expected win above 1 through 4 weeks, but in the real standings they still sit above the GMO's who should break winless streak soon given the respectable 1.42 expected wins that the model gives them. 

```{r expected_wins, echo=FALSE,warning=FALSE,fig.width = 10,fig.height=8,message=FALSE}
kfl_wins_over_expected<-kfl_2021_weekly_data%>%select(win_prob,win,franchise_id,franchise_name,logo,franchise_abbrev)%>%group_by(franchise_id,franchise_name,franchise_abbrev,logo)%>%summarise(expected_wins = sum(win_prob),wins = sum(win))%>%mutate(wins_over_expected = wins - expected_wins)%>%arrange(expected_wins)%>%na.omit()

week_4_table<-kfl_wins_over_expected%>%select(franchise_id,franchise_name,franchise_abbrev,logo,wins,expected_wins,wins_over_expected)%>%mutate(expected_wins = round(expected_wins,2), wins_over_expected = round(wins_over_expected,2))%>%arrange(desc(expected_wins))

week_4_table<-as.data.frame(week_4_table[,c(2:7)])%>%na.omit()

asp_ratio <- 1.618 


################### UPDATE PLOT TITLE ##############################
week4_exp_wins<-ggplot(kfl_wins_over_expected,aes(x=reorder(franchise_abbrev,expected_wins),y=expected_wins))+geom_bar(stat="identity")+geom_image(aes(x=franchise_abbrev,y=expected_wins,image=logo), size = 0.05, by = "width", asp = asp_ratio
)+coord_flip()+ggtitle("Expected Wins through Week 4") + xlab("franchise") + ylab("expected wins")+theme(axis.title.x = element_text(size=14),axis.title.y=element_text(size=14),axis.text = element_text(size = 12),plot.title = element_text(size = 18))

week4_exp_wins

################### UPDATE TABLE TITLE ##############################
week_4_table%>%select(franchise_name,logo,expected_wins,wins_over_expected)%>%gt()%>%gt_img_rows(logo) %>%tab_header(title ="Expected Wins through Week 4")%>%
  cols_label(franchise_name="Team Name",logo="",expected_wins="Expected Wins",wins_over_expected="Wins Over Expected")


#theme(aspect.ratio = 1/asp_ratio)

# include aspect ratio in ggsave
#ggsave(
#  "week4_exp_wins.png", week4_exp_wins, 
#  # make the width equivalent to the aspect.ratio
#  height = 20, width = 20 * asp_ratio, dpi = "retina"
#)






```
## Standings through Week 4

We have clusters at the top and bottom with 5 3-1 teams and 5 1-3. The coming weeks will separate the pretenders from the contenders.

```{r divison records, echo=FALSE,warning=FALSE,fig.width = 10,fig.height=8,message=FALSE}

kfl_standings_wk4<-kfl_2021_weekly_data%>%filter(week<=week_num)

kfl_standings_wk4[41,9]<-0
kfl_standings_wk4[45,9]<-1


franchise_id<-1:12
division_name<-c("Ohio","Glum","Ohio","Indiana","Glum","Ohio","Ohio","Indiana","Indiana","Glum","Glum","Indiana")
kfl_division<-as.data.frame(cbind(franchise_id,division_name))
kfl_division$franchise_id<-as.numeric(kfl_division$franchise_id)


kfl_division_data<- sqldf("Select A.*,B.division_name
      from kfl_standings_wk4 A
      left outer join kfl_division B
      on A.Franchise_ID = B.Franchise_ID
      ")

team_df<-kfl_division_data%>%filter(week == 1)%>%select(franchise_id,franchise_name,logo,division_name)

joined_df<-kfl_division_data%>%select(franchise_id,franchise_name,logo,division_name,win)%>%group_by(franchise_name,franchise_id)%>%summarise(
         Wins = length(win[win==1]),
         Losses = length(win[win==0]),
         outcomes = list(win), .groups = "drop") %>% 
     left_join(team_df, by = c("franchise_id" = "franchise_id","franchise_name"="franchise_name")) %>% 
     select(logo,franchise_name, division_name, Wins:outcomes)
 

###################### NEED TO UPDATE THE TITLE HEADER FOR WEEK  ##########################
standings_df<-joined_df%>%select(franchise_name,logo,division_name,Wins,Losses,outcomes)%>%arrange(desc(Wins))%>%gt(groupname_col="division_name")%>%gt_plt_winloss(outcomes,max_wins = 14)%>%gt_img_rows(logo)%>%tab_header(title ="Standings Through Week 4")%>%cols_label(franchise_name="Team",logo="",outcomes="Win-Loss Streak")

standings_df

```

### Best Ball Week 4 Update

 Jeff's Slugs are squeezing the most out of their roster with only 16 points left on the bench through 4 weeks. Michaeal's GMOs are doing the best with what they have but sadly they don't have much and look to another season in the basement of the KFL. They are our only team who would have not won a game in a best ball format. At the bottom of the best ball standings we have the Tenderloin's and the Sunday Winners who have left over 100 points on the bench each and a shocker that even with Nick's (Un)Lucky Ducks leaving tons of points on the bench they are still the only undefeated franchise in the league. 
 
 Though Elliot's Last Good Tight End took their first L this week, they continue to show they have talent and lead the league in best ball points with 460. 

```{r bench depth, echo=FALSE,warning=FALSE,fig.width = 10,fig.height=8,message=FALSE}

bb_results<-data.frame()
for (i in 1:week_num){
 player_rankings<-starters %>%filter(week==i)%>%group_by(franchise_name,pos) %>% mutate(rank = rank(desc(player_score),ties.method="random"),pos_rank=paste0(pos,"_",rank))%>%arrange(franchise_name,pos,desc(player_score))

 
optimal_starters<-player_rankings%>%filter(pos_rank=='DST_1'|pos_rank=='K_1'|pos_rank=='QB_1'|pos_rank=='RB_1'|pos_rank=='WR_1'|pos_rank=='WR_2'|pos_rank=='TE_1')

flex_spot<-player_rankings%>%filter(pos_rank=='RB_2'|pos_rank=='WR_3')%>%group_by(franchise_name)%>%mutate(flex = ifelse(pos_rank))%>%arrange(franchise_name,desc(player_score))%>%filter(row_number()==1)

best_ball_team <-optimal_starters%>%union_all(flex_spot)

bb_team <-best_ball_team%>%select(franchise_id,franchise_name,pos_rank,player_score)%>%group_by(franchise_id,franchise_name,pos_rank,player_score)%>%distinct()

bb_team<-bb_team%>%group_by(franchise_id,franchise_name)%>%mutate(player_n = n(),team_score = sum(player_score))%>%select(franchise_name,franchise_id,player_n,team_score)%>%distinct()

actual_result<-ff_schedule(espn_conn_2022)%>%select(week,franchise_id,franchise_score)%>%filter(week==i)

bb_team <-bb_team%>%inner_join(actual_result,by=c("franchise_id"="franchise_id"))%>%mutate(pts_below_optimal = franchise_score - team_score)%>%arrange(desc(pts_below_optimal))

bb_team<-as.data.frame(bb_team%>%inner_join(kfl_franchises_2021,by=c("franchise_id"="franchise_id"))%>%select(franchise_name.x,week,logo,player_n,team_score,franchise_score,pts_below_optimal)%>%select(-franchise_id))

bb_schedule<-ff_schedule(espn_conn_2022)%>%filter(week ==i)%>%inner_join(bb_team%>%dplyr::select(franchise_id,team_score),by=c("franchise_id"="franchise_id"))%>%inner_join(bb_team%>%dplyr::select(franchise_id,team_score),by=c("opponent_id"="franchise_id"))%>%rename(fran_bb_score = team_score.x, opp_bb_score = team_score.y)%>%mutate(bb_win = ifelse(fran_bb_score>opp_bb_score,1,0))

bb_team<-bb_team%>%inner_join(bb_schedule%>%dplyr::select(franchise_id,bb_win),by=c("franchise_id"="franchise_id"))%>%mutate(bb_loss=1-bb_win)


bb_results<-rbind(bb_results,bb_team)
}

bb_team<-bb_results%>%
  select(-week,-player_n)%>%
group_by(franchise_id,franchise_name.x,logo)%>%
  mutate(total_bb_score =sum(team_score),
         total_actual_score=sum(franchise_score),
         pts_below_optimal=sum(pts_below_optimal),
         bb_wins = sum(bb_win),
         bb_losses = sum(bb_loss))%>%select(-franchise_id,-bb_win,-bb_loss,-franchise_score,-team_score)%>%distinct()%>%arrange(desc(pts_below_optimal))

bb_team<-as.data.frame(bb_team)
```


```{r bench, echo=FALSE,warning=FALSE,fig.width = 10,fig.height=8,message=FALSE}

bb_table <- bb_team%>%gt()%>%
  text_transform(
      locations = cells_body(
        vars(logo)
      ),
      fn = function(x) {
        web_image(
          url = x,
          height = 25
        )
      })%>%
   cols_label(
      franchise_name.x = "KFL Team",
      logo = "logo",
      total_bb_score = "Total Score - Optimal Lineup",
      total_actual_score = "Actual Total Score",
      pts_below_optimal = "Points under Optimal",
      bb_wins = "Best Ball Wins",
      bb_losses = "Best Ball Loss"
    )

bb_table
```

## Stud of the Week

Our first Tight End as stud of the week. T.J. Hockenson took advantage of the Sun God Am'ron St. Brown's absence to put up a massive 31 points which was 25 over the expected 6 points for a tight end. Scott's Titans reaped the rewards of this performance to make up for Brown's absence who is also on the Titans. I am sure this is the high point for Hockenson's season so he should enjoy being the KFL stud of the week. Even with the Stud performance though Hockenson couldn't lead his team above the Kraken losing a heartbreaking tie.

```{r stud, echo=FALSE,warning=FALSE,fig.width = 10,fig.height=8,message=FALSE}
team_df <- espnscrapeR::get_nfl_teams() %>% 
  select(team = team_abb, logo)

exclude_pos <-c('BE','DST')

stud<-starters%>%filter(lineup_slot != exclude_pos & week==week_num)%>%mutate(proj_diff=player_score-projected_score)%>%slice_max(proj_diff)%>%select(franchise_name,player_name,team,player_score,projected_score,proj_diff)
dud<-starters%>%filter(lineup_slot != exclude_pos& week==week_num)%>%mutate(proj_diff=player_score-projected_score)%>%slice_min(proj_diff)%>%select(franchise_name,player_name,team,player_score,projected_score,proj_diff)


dud<-dud%>%inner_join(team_df,by=c('team'='team'))
stud<-stud%>%inner_join(team_df,by=c('team'='team'))

stud_table <- stud%>%select(franchise_name,player_name,logo,team,player_score,projected_score,proj_diff) %>%
  gt() %>%
  text_transform(
      locations = cells_body(
        vars(logo)
      ),
      fn = function(x) {
        web_image(
          url = x,
          height = 25
        )
      })%>%
   cols_label(
      franchise_name = "KFL Team",
      player_name = "player",
      logo = "",
      team = "Team",
      player_score = "actual score",
      projected_score = "projection",
      proj_diff = "difference"
    )
stud_table

```
## Dud of the Week

Sadly, we have to give Tua the dud of the week putting up a measly 2 points after his horrific injury. Hopefully, Tua will make a comeback because he was playing well prior to the concussions.

```{r dud, echo=FALSE,warning=FALSE,fig.width = 10,fig.height=8,message=FALSE}
dud_table <- dud%>%select(franchise_name,player_name,logo,team,player_score,projected_score,proj_diff) %>%
  gt() %>%
  text_transform(
      locations = cells_body(
        vars(logo)
      ),
      fn = function(x) {
        web_image(
          url = x,
          height = 25
        )
      })%>%
   cols_label(
      franchise_name = "KFL Team",
      player_name = "player",
      logo = "",
      team = "Team",
      player_score = "actual score",
      projected_score = "projection",
      proj_diff = "difference"
    )
dud_table

```