---
title: "Week 3 2022 Recap"
author: "Samuel Kirschner"
date: "2022-09-26T13:09:13-06:00"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ffscrapr)
library(dplyr)
library(ggplot2)
library(gt)
library(sqldf)
library(gtExtras)
library(tidymodels)
library(tidyr)
library(DBI)
library(ggimage)
#remotes::install_github("jthomasmock/espnscrapeR")
library(espnscrapeR)



#### INPUT WHAT WEEK IT IS HERE ####
week_num<-3

swid<-"{FB82F2B3-89EB-4E1A-9677-B1333054B883}"

ESPN_S2<-"AECkrDCmS9dG4sssE55%2BNQRasW7jwjJVFyEUw6mR2wLFylIkvGy89sVJxc8GMvORMKJaBksVZfGnYB4BXKiM80pkw352zpsk%2BEbMg2ILadghIeMp2iz1Etpke3FMM2pyVdNg23edB2DoYIHMoUT72KS3Ho9bKNLPkkJ761JMh4o4ikTDTeL1uBJYNiahAMg1uvT4mHvszAP1UoDHQGLUPuOLchpl8Os8BsibtvBOTn778ksyx6gU513HRQGrpge7iQ9G04xOeWwYAjXfwcxa%2FCQr3Gx9kECPmidN9%2F7PrHImFQ%3D%3D"

espn_conn_2022 <- espn_connect(
  season = 2022,
  league_id = 214888,
  espn_s2 = ESPN_S2,
  swid = swid
)

kfl_schedule_2021<-ff_schedule(espn_conn_2022)
kfl_pscores<-ff_playerscores(espn_conn_2022)
starters<-ff_starters(espn_conn_2022)


kfl_franchises_2021<-ff_franchises(espn_conn_2022)
#kfl_franchises_2021$logo[10]<-"https://image.shutterstock.com/image-vector/lightning-warrior-logo-template-design-260nw-1019248687.jpg"
kfl_franchises_2021$logo[3]<-"https://www.pngkit.com/png/detail/269-2691865_winners-logo-winners.png"

kfl_db <-"kfl_ff_data" 
host_db <- "localhost"
db_port <- "5432"
db_user <- "postgres"  
db_password <- "Cincinnati1"
kfl_con <- dbConnect(RPostgres::Postgres(), dbname = kfl_db, host=host_db, port=db_port, user=db_user, password=db_password)   

kfl_schedule<-dbReadTable(kfl_con,"kfl_schedule")
kfl_franchises<-dbReadTable(kfl_con,"kfl_franchise")

```



```{r wins over expected model run, echo=FALSE}



score_distribution_data<-sqldf(
"select a.franchise_id,b.user_name,a.Years,a.week,franchise_score as score_for, opponent_score as score_against
 from kfl_schedule a
 left outer join kfl_franchises b 
 on a.franchise_id = b.franchise_id and a.Years = b.Years
 where week < 15"
)

score_distribution_data<-score_distribution_data%>%unite(YR_WK, c("Years","week"),remove = FALSE)

score_distribution_data<-score_distribution_data%>%group_by(YR_WK)%>%mutate(avg_weekly_score=mean(score_for),win = case_when(score_for > score_against ~ 1,score_for < score_against ~ 0))

score_distribution_data$win<-as.factor(score_distribution_data$win)

fitted_logistic_model<- logistic_reg() %>%
        # Set the engine
        set_engine("glm") %>%
        # Set the mode
        set_mode("classification") %>%
        # Fit the model
        fit(win~score_for, data = score_distribution_data)

```

```{r wins over expected model, echo=FALSE,warning=FALSE}
## 2021 win probs

kfl_2021_win_prob_data<-kfl_schedule_2021%>%rename( score_for = franchise_score)%>%filter(week<=week_num)

pred_prob_2021 <- predict(fitted_logistic_model,
                          new_data = kfl_2021_win_prob_data,
                          type = "prob")

colnames(pred_prob_2021)<-c("loss_prob", "win_prob")

kfl_2021_win_prob_data<-cbind(kfl_2021_win_prob_data,pred_prob_2021)

kfl_2021_win_prob_data<-kfl_2021_win_prob_data%>%mutate(win = case_when(score_for > opponent_score ~ 1,score_for < opponent_score ~ 0))

kfl_2021_win_prob_data[37,9]<-1
kfl_2021_win_prob_data[40,9]<-0
kfl_2021_win_prob_data[39,9]<-1
kfl_2021_win_prob_data[44,9]<-0


kfl_2021_weekly_data<-sqldf("SELECT A.*,B.franchise_name,B.franchise_abbrev,B.logo,C.franchise_name as opponent_franchise,C.franchise_abbrev as opponent_abbrev,C.logo as opponent_logo
      FROM kfl_2021_win_prob_data A
      LEFT OUTER JOIN kfl_franchises_2021 B
      on a.franchise_id = b.franchise_id
      LEFT OUTER JOIN kfl_franchises_2021 C
      on a.opponent_id = c.franchise_id")
```
## Week 3 Recap

We have two remaining undefeated teams after three weeks, one deserving and the other not so much. Elliot's Last Good Tight End look to ride the Lamar Jackson - Mark Andrews combo to KFL glory, while the (Un)Lucky Ducks have been enjoying a cruise in a cupcake division. They win yet again with a pathetic 62 points and will surely come down to earth soon. The Ohio division looks to be another bloodbath of mediocrity as the Monarchs got a much deserved win and another low scoring battle saw the Slugs take the divison lead.

Meanwhile the Tenderloin's got their first win of the season and we have a pair of winless franchises; The Supermarket Duds and Michael's GMOs who both had ok days, but couldn't find a win. 

```{r wk4 review, echo=FALSE,warning=FALSE,,message=FALSE}

wk4_matchups<-kfl_2021_weekly_data%>%filter(week==week_num &win==1)


############################## UPDATE THE WEEK #####################################################
wk4_matchups<-sqldf("select A.*,B.win_prob as opp_win_prob
                    FROM wk4_matchups A
                    LEFT OUTER JOIN kfl_2021_weekly_data B
                    ON A.opponent_id=B.franchise_id
                    WHERE B.week = 3")

wk4_matchups<-wk4_matchups%>%select(franchise_name,logo,win_prob,score_for,opponent_score,opp_win_prob,opponent_logo,opponent_franchise)


########################## UPDATE THE TITLE OF THE TAB HEADER #######################################
wk4_matchups%>%mutate(win_prob=round(win_prob,2),opp_win_prob=round(opp_win_prob,2))%>%gt()%>%gt_img_rows(logo)%>%gt_img_rows(opponent_logo)%>%
  tab_header(title ="Week 3 2022 matchups")%>%
  cols_label(franchise_name="Winning Team",logo="",win_prob= "Win Probability",score_for="Score",opponent_score="Score",opp_win_prob="Win Probability",opponent_logo="",opponent_franchise="Losing Team")%>%
  tab_style(
    style = list(
      cell_borders(
        sides = "left",
        color = "black",
        weight = px(3)
      )
    ),
    locations = list(
      cells_body(
        columns = vars(opponent_score)
      )
    )
  ) %>%
  # We use tab_style() to change style of cells
  # cell_borders() provides the formatting
  # locations tells it where
  # Add black borders to the bottom of all the column labels
  tab_style(
    style = list(
      cell_borders(
        sides = "bottom",
        color = "black",
        weight = px(3)
      )
    ),
    locations = list(
      cells_column_labels(
        columns = gt::everything()
      )
    )
  )%>%
   data_color(
    columns = vars(score_for),
    colors = scales::col_numeric(
      palette = c("#c9ecb4"),
      domain = NULL
    ))%>%
   data_color(
    columns = vars(win_prob),
    colors = scales::col_numeric(
      palette = c("#ffffff", "#f2fbd2", "#c9ecb4", "#93d3ab", "#35b0ab"),
      domain = NULL
    ))%>%
   data_color(
    columns = vars(opponent_score),
    colors = scales::col_numeric(
      palette = c("#EF9A9AFF"),
      domain = NULL
    ))%>%
   data_color(
    columns = vars(opp_win_prob),
    colors = scales::col_numeric(
      # Using a function from paletteer to generate a vector of colors
      # Note that you need to wrap paletteer_d outputs in as.character()
      palette = as.character(paletteer::paletteer_d("ggsci::red_material", n = 5)),
      # Domain is from scales::col_numeric
      # Domain = The possible values that can be mapped
      # We don't HAVE to set a range for this since
      # we're using it inside data_color()
      domain = NULL
    ))

```

## Expected Wins power rankings through week 3 2022

Elliot's Last Good Tight End cruise to a commanding and comfortable lead of our power rankings. The Onionhead's follow with the Monarchs and Slugs close behind. The power rankings suggest that the winless GMO's are due a win soon and currently the unluckiest franchise in the league. Meanwhile the (UN)Lucky Ducks are already at a ridciulous 2.3 wins over expected putting up the 3rd lowest expected wins this year, but remaining undefeated.


```{r expected_wins, echo=FALSE,warning=FALSE,fig.width = 10,fig.height=8,message=FALSE}
kfl_wins_over_expected<-kfl_2021_weekly_data%>%select(win_prob,win,franchise_id,franchise_name,logo,franchise_abbrev)%>%group_by(franchise_id,franchise_name,franchise_abbrev,logo)%>%summarise(expected_wins = sum(win_prob),wins = sum(win))%>%mutate(wins_over_expected = wins - expected_wins)%>%arrange(expected_wins)%>%na.omit()

week_4_table<-kfl_wins_over_expected%>%select(franchise_id,franchise_name,franchise_abbrev,logo,wins,expected_wins,wins_over_expected)%>%mutate(expected_wins = round(expected_wins,2), wins_over_expected = round(wins_over_expected,2))%>%arrange(desc(expected_wins))

week_4_table<-as.data.frame(week_4_table[,c(2:7)])%>%na.omit()

asp_ratio <- 1.618 


################### UPDATE PLOT TITLE ##############################
week4_exp_wins<-ggplot(kfl_wins_over_expected,aes(x=reorder(franchise_abbrev,expected_wins),y=expected_wins))+geom_bar(stat="identity")+geom_image(aes(x=franchise_abbrev,y=expected_wins,image=logo), size = 0.05, by = "width", asp = asp_ratio
)+coord_flip()+ggtitle("Expected Wins through Week 3") + xlab("franchise") + ylab("expected wins")+theme(axis.title.x = element_text(size=14),axis.title.y=element_text(size=14),axis.text = element_text(size = 12),plot.title = element_text(size = 18))

week4_exp_wins

################### UPDATE TABLE TITLE ##############################
week_4_table%>%select(franchise_name,logo,expected_wins,wins_over_expected)%>%gt()%>%gt_img_rows(logo) %>%tab_header(title ="Expected Wins through Week 3")%>%
  cols_label(franchise_name="Team Name",logo="",expected_wins="Expected Wins",wins_over_expected="Wins Over Expected")


#theme(aspect.ratio = 1/asp_ratio)

# include aspect ratio in ggsave
#ggsave(
#  "week4_exp_wins.png", week4_exp_wins, 
#  # make the width equivalent to the aspect.ratio
#  height = 20, width = 20 * asp_ratio, dpi = "retina"
#)






```
## Standings through Week 3



```{r divison records, echo=FALSE,warning=FALSE,fig.width = 10,fig.height=8,message=FALSE}

kfl_standings_wk4<-kfl_2021_weekly_data%>%filter(week<=week_num)

franchise_id<-1:12
division_name<-c("Ohio","Glum","Ohio","Indiana","Glum","Ohio","Ohio","Indiana","Indiana","Glum","Glum","Indiana")
kfl_division<-as.data.frame(cbind(franchise_id,division_name))
kfl_division$franchise_id<-as.numeric(kfl_division$franchise_id)


kfl_division_data<- sqldf("Select A.*,B.division_name
      from kfl_standings_wk4 A
      left outer join kfl_division B
      on A.Franchise_ID = B.Franchise_ID
      ")

team_df<-kfl_division_data%>%filter(week == 1)%>%select(franchise_id,franchise_name,logo,division_name)

joined_df<-kfl_division_data%>%select(franchise_id,franchise_name,logo,division_name,win)%>%group_by(franchise_name,franchise_id)%>%summarise(
         Wins = length(win[win==1]),
         Losses = length(win[win==0]),
         outcomes = list(win), .groups = "drop") %>% 
     left_join(team_df, by = c("franchise_id" = "franchise_id","franchise_name"="franchise_name")) %>% 
     select(logo,franchise_name, division_name, Wins:outcomes)
 

###################### NEED TO UPDATE THE TITLE HEADER FOR WEEK  ##########################
standings_df<-joined_df%>%select(franchise_name,logo,division_name,Wins,Losses,outcomes)%>%arrange(desc(Wins))%>%gt(groupname_col="division_name")%>%gt_plt_winloss(outcomes,max_wins = 14)%>%gt_img_rows(logo)%>%tab_header(title ="Standings Through Week 3")%>%cols_label(franchise_name="Team",logo="",outcomes="Win-Loss Streak")

standings_df

```

### Best Ball Week 3 Update

After 3 weeks, Jeff's Slugs look to be getting the most out of their team and putting out a really optimal lineup leading to his 2-1 record matching his 2-1 best ball record. Not surprising, Elliot's Last Good Tight End leads in points and also has points to spare on the bench. Poor Kurt's Tenderloins would be sitting at 3rd in Best Ball overall points, but just has had shambolic time managing the franchise. Finally, it is important to note that the 2-1 Bluecats sit at 0-3 based on best ball record and don't leave a lot of points on the bench suggesting they will really struggle from here on out and likely miss the playoffs once again. 

```{r bench depth, echo=FALSE,warning=FALSE,fig.width = 10,fig.height=8,message=FALSE}

bb_results<-data.frame()
for (i in 1:week_num){
 player_rankings<-starters %>%filter(week==i)%>%group_by(franchise_name,pos) %>% mutate(rank = rank(desc(player_score),ties.method="random"),pos_rank=paste0(pos,"_",rank))%>%arrange(franchise_name,pos,desc(player_score))

 
optimal_starters<-player_rankings%>%filter(pos_rank=='DST_1'|pos_rank=='K_1'|pos_rank=='QB_1'|pos_rank=='RB_1'|pos_rank=='WR_1'|pos_rank=='WR_2'|pos_rank=='TE_1')

flex_spot<-player_rankings%>%filter(pos_rank=='RB_2'|pos_rank=='WR_3')%>%group_by(franchise_name)%>%mutate(flex = ifelse(pos_rank))%>%arrange(franchise_name,desc(player_score))%>%filter(row_number()==1)

best_ball_team <-optimal_starters%>%union_all(flex_spot)

bb_team <-best_ball_team%>%select(franchise_id,franchise_name,pos_rank,player_score)%>%group_by(franchise_id,franchise_name,pos_rank,player_score)%>%distinct()

bb_team<-bb_team%>%group_by(franchise_id,franchise_name)%>%mutate(player_n = n(),team_score = sum(player_score))%>%select(franchise_name,franchise_id,player_n,team_score)%>%distinct()

actual_result<-ff_schedule(espn_conn_2022)%>%select(week,franchise_id,franchise_score)%>%filter(week==i)

bb_team <-bb_team%>%inner_join(actual_result,by=c("franchise_id"="franchise_id"))%>%mutate(pts_below_optimal = franchise_score - team_score)%>%arrange(desc(pts_below_optimal))

bb_team<-as.data.frame(bb_team%>%inner_join(kfl_franchises_2021,by=c("franchise_id"="franchise_id"))%>%select(franchise_name.x,week,logo,player_n,team_score,franchise_score,pts_below_optimal)%>%select(-franchise_id))

bb_schedule<-ff_schedule(espn_conn_2022)%>%filter(week ==i)%>%inner_join(bb_team%>%dplyr::select(franchise_id,team_score),by=c("franchise_id"="franchise_id"))%>%inner_join(bb_team%>%dplyr::select(franchise_id,team_score),by=c("opponent_id"="franchise_id"))%>%rename(fran_bb_score = team_score.x, opp_bb_score = team_score.y)%>%mutate(bb_win = ifelse(fran_bb_score>opp_bb_score,1,0))

bb_team<-bb_team%>%inner_join(bb_schedule%>%dplyr::select(franchise_id,bb_win),by=c("franchise_id"="franchise_id"))%>%mutate(bb_loss=1-bb_win)


bb_results<-rbind(bb_results,bb_team)
}

bb_team<-bb_results%>%
  select(-week,-player_n)%>%
group_by(franchise_id,franchise_name.x,logo)%>%
  mutate(total_bb_score =sum(team_score),
         total_actual_score=sum(franchise_score),
         pts_below_optimal=sum(pts_below_optimal),
         bb_wins = sum(bb_win),
         bb_losses = sum(bb_loss))%>%select(-franchise_id,-bb_win,-bb_loss,-franchise_score,-team_score)%>%distinct()%>%arrange(desc(pts_below_optimal))

bb_team<-as.data.frame(bb_team)
```


```{r bench, echo=FALSE,warning=FALSE,fig.width = 10,fig.height=8,message=FALSE}

bb_table <- bb_team%>%gt()%>%
  text_transform(
      locations = cells_body(
        vars(logo)
      ),
      fn = function(x) {
        web_image(
          url = x,
          height = 25
        )
      })%>%
   cols_label(
      franchise_name.x = "KFL Team",
      logo = "logo",
      total_bb_score = "Total Score - Optimal Lineup",
      total_actual_score = "Actual Total Score",
      pts_below_optimal = "Points under Optimal",
      bb_wins = "Best Ball Wins",
      bb_losses = "Best Ball Loss"
    )

bb_table
```

## Stud of the Week

Yet again Elliot's Last Good Tight End get another massive week from Lamar Jackson who is once again looking in MVP form. If he keeps putting up these types of results this will be Elliot's title to lose.

```{r stud, echo=FALSE,warning=FALSE,fig.width = 10,fig.height=8,message=FALSE}
team_df <- espnscrapeR::get_nfl_teams() %>% 
  select(team = team_abb, logo)

exclude_pos <-c('BE','DST')

stud<-starters%>%filter(lineup_slot != exclude_pos & week==week_num)%>%mutate(proj_diff=player_score-projected_score)%>%slice_max(proj_diff)%>%select(franchise_name,player_name,team,player_score,projected_score,proj_diff)
dud<-starters%>%filter(lineup_slot != exclude_pos& week==week_num)%>%mutate(proj_diff=player_score-projected_score)%>%slice_min(proj_diff)%>%select(franchise_name,player_name,team,player_score,projected_score,proj_diff)


dud<-dud%>%inner_join(team_df,by=c('team'='team'))
stud<-stud%>%inner_join(team_df,by=c('team'='team'))

stud_table <- stud%>%select(franchise_name,player_name,logo,team,player_score,projected_score,proj_diff) %>%
  gt() %>%
  text_transform(
      locations = cells_body(
        vars(logo)
      ),
      fn = function(x) {
        web_image(
          url = x,
          height = 25
        )
      })%>%
   cols_label(
      franchise_name = "KFL Team",
      player_name = "player",
      logo = "",
      team = "Team",
      player_score = "actual score",
      projected_score = "projection",
      proj_diff = "difference"
    )
stud_table

```
## Dud of the Week

The dud this week goes to last year's fraud of a Super Bowl champ, Matthew Stafford. His interception happy ways led him to put up a measly 9 points when he was projected for 22 and was a big reason why Alex's Kraken fell by a measly three points to Nick's (Un)Lucky Ducks.


```{r dud, echo=FALSE,warning=FALSE,fig.width = 10,fig.height=8,message=FALSE}
dud_table <- dud%>%select(franchise_name,player_name,logo,team,player_score,projected_score,proj_diff) %>%
  gt() %>%
  text_transform(
      locations = cells_body(
        vars(logo)
      ),
      fn = function(x) {
        web_image(
          url = x,
          height = 25
        )
      })%>%
   cols_label(
      franchise_name = "KFL Team",
      player_name = "player",
      logo = "",
      team = "Team",
      player_score = "actual score",
      projected_score = "projection",
      proj_diff = "difference"
    )
dud_table

```