---
title: "Week 5 2021 Recap"
author: "Sam Kirschner"
date: "2021-10-06T13:09:13-06:00"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ffscrapr)
library(dplyr)
library(ggplot2)
library(gt)
library(sqldf)
library(gtExtras)
library(tidymodels)
library(tidyr)
library(DBI)
library(ggimage)



#### INPUT WHAT WEEK IT IS HERE ####
week_num<-5

swid<-"{FB82F2B3-89EB-4E1A-9677-B1333054B883}"

ESPN_S2<-"AECkrDCmS9dG4sssE55%2BNQRasW7jwjJVFyEUw6mR2wLFylIkvGy89sVJxc8GMvORMKJaBksVZfGnYB4BXKiM80pkw352zpsk%2BEbMg2ILadghIeMp2iz1Etpke3FMM2pyVdNg23edB2DoYIHMoUT72KS3Ho9bKNLPkkJ761JMh4o4ikTDTeL1uBJYNiahAMg1uvT4mHvszAP1UoDHQGLUPuOLchpl8Os8BsibtvBOTn778ksyx6gU513HRQGrpge7iQ9G04xOeWwYAjXfwcxa%2FCQr3Gx9kECPmidN9%2F7PrHImFQ%3D%3D"

espn_conn_2021 <- espn_connect(
  season = 2021,
  league_id = 214888,
  espn_s2 = ESPN_S2,
  swid = swid
)

kfl_schedule_2021<-ff_schedule(espn_conn_2021)


kfl_franchises_2021<-ff_franchises(espn_conn_2021)
kfl_franchises_2021$logo[10]<-"https://image.shutterstock.com/image-vector/lightning-warrior-logo-template-design-260nw-1019248687.jpg"
kfl_franchises_2021$logo[3]<-"https://www.pngkit.com/png/detail/269-2691865_winners-logo-winners.png"

kfl_db <-"kfl_ff_data" 
host_db <- "localhost"
db_port <- "5432"
db_user <- "postgres"  
db_password <- "Cincinnati1"
kfl_con <- dbConnect(RPostgres::Postgres(), dbname = kfl_db, host=host_db, port=db_port, user=db_user, password=db_password)   

kfl_schedule<-dbReadTable(kfl_con,"kfl_schedule")
kfl_franchises<-dbReadTable(kfl_con,"kfl_franchises")

```

## Week 5 Recap: Monday Night Miracle for the Unlucky Ducks

A massive performance from Lamar Jackson and Mark Andrews helped the once Fightin' Ducks slide past the Slugs in a 148-120 slug fest. The Slug's certainly can count themselves unlucky after putting up 120, but losing out thanks to Lamar Jackson's masterclass second half performance. 

Speaking of luck, the Bluecat's faced a different type of luck or lack thereof, losing Russell Wilson on Thursday night football mid game. This handicap the team and thanks to Dawson Knox's 17 point game compared to Darren Waller's pitiful performance, the Titans squeaked by in a 79-77 game. 

The Sunday Winner's are back on track with a 2-3 season defeating the Kraken who fall to 3-2. The Sunday Winner's are trying to do it the hard way getting no points from their running backs and still pulling out a W. There is a reason Brady is the GOAT. The Tenderloins and SWAB trounced their respective opponents who both have a lot to be GLUM about putting up poor point totals and both at the bottom of the expected wins table.
```{r wins over expected model run, echo=FALSE}



score_distribution_data<-sqldf(
"select a.franchise_id,b.user_name,a.Years,a.week,franchise_score as score_for, opponent_score as score_against
 from kfl_schedule a
 left outer join kfl_franchises b 
 on a.franchise_id = b.franchise_id and a.Years = b.Years
 where week < 15"
)

score_distribution_data<-score_distribution_data%>%unite(YR_WK, c("Years","week"),remove = FALSE)

score_distribution_data<-score_distribution_data%>%group_by(YR_WK)%>%mutate(avg_weekly_score=mean(score_for),win = case_when(score_for > score_against ~ 1,score_for < score_against ~ 0))

score_distribution_data$win<-as.factor(score_distribution_data$win)

fitted_logistic_model<- logistic_reg() %>%
        # Set the engine
        set_engine("glm") %>%
        # Set the mode
        set_mode("classification") %>%
        # Fit the model
        fit(win~score_for, data = score_distribution_data)

```

```{r wins over expected model, echo=FALSE,warning=FALSE}
## 2021 win probs

kfl_2021_win_prob_data<-kfl_schedule_2021%>%rename( score_for = franchise_score)%>%filter(week<=week_num)

pred_prob_2021 <- predict(fitted_logistic_model,
                          new_data = kfl_2021_win_prob_data,
                          type = "prob")

colnames(pred_prob_2021)<-c("loss_prob", "win_prob")

kfl_2021_win_prob_data<-cbind(kfl_2021_win_prob_data,pred_prob_2021)

kfl_2021_win_prob_data<-kfl_2021_win_prob_data%>%mutate(win = case_when(score_for > opponent_score ~ 1,score_for < opponent_score ~ 0))

kfl_2021_win_prob_data[37,9]<-1
kfl_2021_win_prob_data[40,9]<-0
kfl_2021_win_prob_data[39,9]<-1
kfl_2021_win_prob_data[44,9]<-0


kfl_2021_weekly_data<-sqldf("SELECT A.*,B.franchise_name,B.franchise_abbrev,B.logo,C.franchise_name as opponent_franchise,C.franchise_abbrev as opponent_abbrev,C.logo as opponent_logo
      FROM kfl_2021_win_prob_data A
      LEFT OUTER JOIN kfl_franchises_2021 B
      on a.franchise_id = b.franchise_id
      LEFT OUTER JOIN kfl_franchises_2021 C
      on a.opponent_id = c.franchise_id")
```

```{r wk4 review, echo=FALSE,warning=FALSE,,message=FALSE}

wk4_matchups<-kfl_2021_weekly_data%>%filter(week==week_num &win==1)


############################## UPDATE THE WEEK #####################################################
wk4_matchups<-sqldf("select A.*,B.win_prob as opp_win_prob
                    FROM wk4_matchups A
                    LEFT OUTER JOIN kfl_2021_weekly_data B
                    ON A.opponent_id=B.franchise_id
                    WHERE B.week = 5")

wk4_matchups<-wk4_matchups%>%select(franchise_name,logo,win_prob,score_for,opponent_score,opp_win_prob,opponent_logo,opponent_franchise)


########################## UPDATE THE TITLE OF THE TAB HEADER #######################################
wk4_matchups%>%mutate(win_prob=round(win_prob,2),opp_win_prob=round(opp_win_prob,2))%>%gt()%>%gt_img_rows(logo)%>%gt_img_rows(opponent_logo)%>%
  tab_header(title ="Week 5 matchups")%>%
  cols_label(franchise_name="Winning Team",logo="",win_prob= "Win Probability",score_for="Score",opponent_score="Score",opp_win_prob="Win Probability",opponent_logo="",opponent_franchise="Losing Team")%>%
  tab_style(
    style = list(
      cell_borders(
        sides = "left",
        color = "black",
        weight = px(3)
      )
    ),
    locations = list(
      cells_body(
        columns = vars(opponent_score)
      )
    )
  ) %>%
  # We use tab_style() to change style of cells
  # cell_borders() provides the formatting
  # locations tells it where
  # Add black borders to the bottom of all the column labels
  tab_style(
    style = list(
      cell_borders(
        sides = "bottom",
        color = "black",
        weight = px(3)
      )
    ),
    locations = list(
      cells_column_labels(
        columns = gt::everything()
      )
    )
  )%>%
   data_color(
    columns = vars(score_for),
    colors = scales::col_numeric(
      palette = c("#c9ecb4"),
      domain = NULL
    ))%>%
   data_color(
    columns = vars(win_prob),
    colors = scales::col_numeric(
      palette = c("#ffffff", "#f2fbd2", "#c9ecb4", "#93d3ab", "#35b0ab"),
      domain = NULL
    ))%>%
   data_color(
    columns = vars(opponent_score),
    colors = scales::col_numeric(
      palette = c("#EF9A9AFF"),
      domain = NULL
    ))%>%
   data_color(
    columns = vars(opp_win_prob),
    colors = scales::col_numeric(
      # Using a function from paletteer to generate a vector of colors
      # Note that you need to wrap paletteer_d outputs in as.character()
      palette = as.character(paletteer::paletteer_d("ggsci::red_material", n = 5)),
      # Domain is from scales::col_numeric
      # Domain = The possible values that can be mapped
      # We don't HAVE to set a range for this since
      # we're using it inside data_color()
      domain = NULL
    ))

```

## Expected Wins power rankings through week 5
The Sunday Winner's continue to lead the league in expected wins, but they now have a few contenders hot on their tail. The Indiana division takes the next three spots with the Tenderloins, Kraken, and Supermarket Studs all vying for division supremacy. 

Jeff's Slugs unlucky loss this week balances out a lucky win earlier in the season and puts them right at 3 wins which is exactly as expected. Nick's unlucky Ducks(1-4) however, still trail their expected wins significantly. The 4-1 SWAB can count their lucky stars to be sitting with a league leading record even though they sit at 8th in expected wins. Expect a regression at some point for the SWAB

The rest of the table is relatively close, except for the Warriors who own a measly 1.6 expected wins to match their 1-4 record.


```{r expected_wins, echo=FALSE,warning=FALSE,fig.width = 10,fig.height=8,message=FALSE}
kfl_wins_over_expected<-kfl_2021_weekly_data%>%select(win_prob,win,franchise_id,franchise_name,logo,franchise_abbrev)%>%group_by(franchise_id,franchise_name,franchise_abbrev,logo)%>%summarise(expected_wins = sum(win_prob),wins = sum(win))%>%mutate(wins_over_expected = expected_wins - wins)%>%arrange(expected_wins)

week_4_table<-kfl_wins_over_expected%>%select(franchise_id,franchise_name,franchise_abbrev,logo,wins,expected_wins,wins_over_expected)%>%mutate(expected_wins = round(expected_wins,1), wins_over_expected = round(wins_over_expected,1))%>%arrange(desc(expected_wins))

week_4_table<-as.data.frame(week_4_table[,c(2:7)])

asp_ratio <- 1.618 


################### UPDATE PLOT TITLE ##############################
week4_exp_wins<-ggplot(kfl_wins_over_expected,aes(x=reorder(franchise_abbrev,expected_wins),y=expected_wins))+geom_bar(stat="identity")+geom_image(aes(x=franchise_abbrev,y=expected_wins,image=logo), size = 0.05, by = "width", asp = asp_ratio
)+coord_flip()+ggtitle("Expected Wins through Week 5") + xlab("franchise") + ylab("expected wins")+theme(axis.title.x = element_text(size=14),axis.title.y=element_text(size=14),axis.text = element_text(size = 12),plot.title = element_text(size = 18))

week4_exp_wins

################### UPDATE TABLE TITLE ##############################
week_4_table%>%select(franchise_name,logo,expected_wins,wins_over_expected)%>%gt()%>%gt_img_rows(logo) %>%tab_header(title ="Expected Wins through Week 5")%>%
  cols_label(franchise_name="Team Name",logo="",expected_wins="Expected Wins",wins_over_expected="Wins Over Expected")


#theme(aspect.ratio = 1/asp_ratio)

# include aspect ratio in ggsave
#ggsave(
 # "week5_exp_wins.png", week4_exp_wins, 
  # make the width equivalent to the aspect.ratio
#  height = 20, width = 20 * asp_ratio, dpi = "retina"
#)






```
## Standings through Week 5

The SWAB will look to stay hot on their 4 game winning streak and douse the newly found hopes of the unlucky ducks, while the Bluecats will look to break their 3 game losing streak against the 3-2 Supermarket Studs.The Kraken get a chance to stop the skid against the basement dwelling Warriors. The Tenderloins and Titans face off as well as the Onionheads vs the Slugs in a battle of 3-2 teams to create some separation in the standings. Finally, to round out the weekend, the GMO's hope to cool off the Sunday Winner's and put them further from playoff contention

```{r divison records, echo=FALSE,warning=FALSE,fig.width = 10,fig.height=8,message=FALSE}

kfl_standings_wk4<-kfl_2021_weekly_data%>%filter(week<=week_num)

franchise_id<-1:12
division_name<-c("Ohio","Glum","Ohio","Indiana","Glum","Ohio","Ohio","Indiana","Indiana","Glum","Glum","Indiana")
kfl_division<-as.data.frame(cbind(franchise_id,division_name))
kfl_division$franchise_id<-as.numeric(kfl_division$franchise_id)


kfl_division_data<- sqldf("Select A.*,B.division_name
      from kfl_standings_wk4 A
      left outer join kfl_division B
      on A.Franchise_ID = B.Franchise_ID
      ")

team_df<-kfl_division_data%>%filter(week == 1)%>%select(franchise_id,franchise_name,logo,division_name)

joined_df<-kfl_division_data%>%select(franchise_id,franchise_name,logo,division_name,win)%>%group_by(franchise_name,franchise_id)%>%summarise(
         Wins = length(win[win==1]),
         Losses = length(win[win==0]),
         outcomes = list(win), .groups = "drop") %>% 
     left_join(team_df, by = c("franchise_id" = "franchise_id","franchise_name"="franchise_name")) %>% 
     select(logo,franchise_name, division_name, Wins:outcomes)
 

###################### NEED TO UPDATE THE TITLE HEADER FOR WEEK  ##########################
standings_df<-joined_df%>%select(franchise_name,logo,division_name,Wins,Losses,outcomes)%>%gt(groupname_col="division_name")%>%gt_plt_winloss(outcomes,max_wins = 14)%>%gt_img_rows(logo)%>%tab_header(title ="Standings Through Week 5")%>%cols_label(franchise_name="Team",logo="",outcomes="Win-Loss Streak")

standings_df

```