---
title: "Week 10 2021 Recap"
author: "Sam Kirschner"
date: 2021-11-16T13:09:13-06:00
output: html_document
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ffscrapr)
library(dplyr)
library(ggplot2)
library(gt)
library(sqldf)
library(gtExtras)
library(tidymodels)
library(tidyr)
library(DBI)
library(ggimage)



#### INPUT WHAT WEEK IT IS HERE ####
week_num<-10

swid<-"{FB82F2B3-89EB-4E1A-9677-B1333054B883}"

ESPN_S2<-"AECkrDCmS9dG4sssE55%2BNQRasW7jwjJVFyEUw6mR2wLFylIkvGy89sVJxc8GMvORMKJaBksVZfGnYB4BXKiM80pkw352zpsk%2BEbMg2ILadghIeMp2iz1Etpke3FMM2pyVdNg23edB2DoYIHMoUT72KS3Ho9bKNLPkkJ761JMh4o4ikTDTeL1uBJYNiahAMg1uvT4mHvszAP1UoDHQGLUPuOLchpl8Os8BsibtvBOTn778ksyx6gU513HRQGrpge7iQ9G04xOeWwYAjXfwcxa%2FCQr3Gx9kECPmidN9%2F7PrHImFQ%3D%3D"

espn_conn_2021 <- espn_connect(
  season = 2021,
  league_id = 214888,
  espn_s2 = ESPN_S2,
  swid = swid
)

kfl_schedule_2021<-ff_schedule(espn_conn_2021)


kfl_franchises_2021<-ff_franchises(espn_conn_2021)
kfl_franchises_2021$logo[10]<-"https://image.shutterstock.com/image-vector/lightning-warrior-logo-template-design-260nw-1019248687.jpg"
kfl_franchises_2021$logo[3]<-"https://www.pngkit.com/png/detail/269-2691865_winners-logo-winners.png"

kfl_db <-"kfl_ff_data" 
host_db <- "localhost"
db_port <- "5432"
db_user <- "postgres"  
db_password <- "Cincinnati1"
kfl_con <- dbConnect(RPostgres::Postgres(), dbname = kfl_db, host=host_db, port=db_port, user=db_user, password=db_password)   

kfl_schedule<-dbReadTable(kfl_con,"kfl_schedule")
kfl_franchises<-dbReadTable(kfl_con,"kfl_franchises")

```



```{r wins over expected model run, echo=FALSE}



score_distribution_data<-sqldf(
"select a.franchise_id,b.user_name,a.Years,a.week,franchise_score as score_for, opponent_score as score_against
 from kfl_schedule a
 left outer join kfl_franchises b 
 on a.franchise_id = b.franchise_id and a.Years = b.Years
 where week < 15"
)

score_distribution_data<-score_distribution_data%>%unite(YR_WK, c("Years","week"),remove = FALSE)

score_distribution_data<-score_distribution_data%>%group_by(YR_WK)%>%mutate(avg_weekly_score=mean(score_for),win = case_when(score_for > score_against ~ 1,score_for < score_against ~ 0))

score_distribution_data$win<-as.factor(score_distribution_data$win)

fitted_logistic_model<- logistic_reg() %>%
        # Set the engine
        set_engine("glm") %>%
        # Set the mode
        set_mode("classification") %>%
        # Fit the model
        fit(win~score_for, data = score_distribution_data)

```

```{r wins over expected model, echo=FALSE,warning=FALSE}
## 2021 win probs

kfl_2021_win_prob_data<-kfl_schedule_2021%>%rename( score_for = franchise_score)%>%filter(week<=week_num)

pred_prob_2021 <- predict(fitted_logistic_model,
                          new_data = kfl_2021_win_prob_data,
                          type = "prob")

colnames(pred_prob_2021)<-c("loss_prob", "win_prob")

kfl_2021_win_prob_data<-cbind(kfl_2021_win_prob_data,pred_prob_2021)

kfl_2021_win_prob_data<-kfl_2021_win_prob_data%>%mutate(win = case_when(score_for > opponent_score ~ 1,score_for < opponent_score ~ 0))

kfl_2021_win_prob_data[37,9]<-1
kfl_2021_win_prob_data[40,9]<-0
kfl_2021_win_prob_data[39,9]<-1
kfl_2021_win_prob_data[44,9]<-0


kfl_2021_weekly_data<-sqldf("SELECT A.*,B.franchise_name,B.franchise_abbrev,B.logo,C.franchise_name as opponent_franchise,C.franchise_abbrev as opponent_abbrev,C.logo as opponent_logo
      FROM kfl_2021_win_prob_data A
      LEFT OUTER JOIN kfl_franchises_2021 B
      on a.franchise_id = b.franchise_id
      LEFT OUTER JOIN kfl_franchises_2021 C
      on a.opponent_id = c.franchise_id")
```

```{r wk4 review, echo=FALSE,warning=FALSE,,message=FALSE}

wk4_matchups<-kfl_2021_weekly_data%>%filter(week==week_num &win==1)


############################## UPDATE THE WEEK #####################################################
wk4_matchups<-sqldf("select A.*,B.win_prob as opp_win_prob
                    FROM wk4_matchups A
                    LEFT OUTER JOIN kfl_2021_weekly_data B
                    ON A.opponent_id=B.franchise_id
                    WHERE B.week = 10")

wk4_matchups<-wk4_matchups%>%select(franchise_name,logo,win_prob,score_for,opponent_score,opp_win_prob,opponent_logo,opponent_franchise)


########################## UPDATE THE TITLE OF THE TAB HEADER #######################################
wk4_matchups%>%mutate(win_prob=round(win_prob,2),opp_win_prob=round(opp_win_prob,2))%>%gt()%>%gt_img_rows(logo)%>%gt_img_rows(opponent_logo)%>%
  tab_header(title ="Week 10 matchups")%>%
  cols_label(franchise_name="Winning Team",logo="",win_prob= "Win Probability",score_for="Score",opponent_score="Score",opp_win_prob="Win Probability",opponent_logo="",opponent_franchise="Losing Team")%>%
  tab_style(
    style = list(
      cell_borders(
        sides = "left",
        color = "black",
        weight = px(3)
      )
    ),
    locations = list(
      cells_body(
        columns = vars(opponent_score)
      )
    )
  ) %>%
  # We use tab_style() to change style of cells
  # cell_borders() provides the formatting
  # locations tells it where
  # Add black borders to the bottom of all the column labels
  tab_style(
    style = list(
      cell_borders(
        sides = "bottom",
        color = "black",
        weight = px(3)
      )
    ),
    locations = list(
      cells_column_labels(
        columns = gt::everything()
      )
    )
  )%>%
   data_color(
    columns = vars(score_for),
    colors = scales::col_numeric(
      palette = c("#c9ecb4"),
      domain = NULL
    ))%>%
   data_color(
    columns = vars(win_prob),
    colors = scales::col_numeric(
      palette = c("#ffffff", "#f2fbd2", "#c9ecb4", "#93d3ab", "#35b0ab"),
      domain = NULL
    ))%>%
   data_color(
    columns = vars(opponent_score),
    colors = scales::col_numeric(
      palette = c("#EF9A9AFF"),
      domain = NULL
    ))%>%
   data_color(
    columns = vars(opp_win_prob),
    colors = scales::col_numeric(
      # Using a function from paletteer to generate a vector of colors
      # Note that you need to wrap paletteer_d outputs in as.character()
      palette = as.character(paletteer::paletteer_d("ggsci::red_material", n = 5)),
      # Domain is from scales::col_numeric
      # Domain = The possible values that can be mapped
      # We don't HAVE to set a range for this since
      # we're using it inside data_color()
      domain = NULL
    ))

```

## Expected Wins power rankings through week 10

In what is becoming a back and forth between the Supermarket Studs and Kraken, the number one spot is once again held by the supermarket studs who beat their rivals the GMO for a 96-89 win. They are at a commanding 6.2 expected wins. The Slugs and Kraken tie at exactly 5.8 expected wins and SWAB keeps the top 4 constant at 5.5 expected wins. One of these teams sadly will be left out of the playoffs and the race to the end of the season will be interesting to see which powerhouse gets left home. The Sunday Winners also sit at 5.5 expected wins followed by the Ducks at 5.1 expected wins to round out the top half. Both teams can count themselves incredibly unlucky this year with -1.5 wins over expected and -3.1 wins over expected, respectively. 

The Titan's continue their march to the Glum division title and pull up to 7th with 4.8 expected wins. The Tenderloins and Bluecats at the next two in our rankings - both teams fought to a pathetic 57-44 scoreline last week which makes the Tenderloins look lucky to sit at 6-4 despite a 4.7 expected wins. The rest of the Glum divsion rounds out the bottom with now a huge gap between the Onionheads in 11th (4.3 expected wins) and the Warriors at the bottom with a measly 3 expected wins. 

```{r expected_wins, echo=FALSE,warning=FALSE,fig.width = 10,fig.height=8,message=FALSE}
kfl_wins_over_expected<-kfl_2021_weekly_data%>%select(win_prob,win,franchise_id,franchise_name,logo,franchise_abbrev)%>%group_by(franchise_id,franchise_name,franchise_abbrev,logo)%>%summarise(expected_wins = sum(win_prob),wins = sum(win))%>%mutate(wins_over_expected = wins - expected_wins)%>%arrange(expected_wins)

week_4_table<-kfl_wins_over_expected%>%select(franchise_id,franchise_name,franchise_abbrev,logo,wins,expected_wins,wins_over_expected)%>%mutate(expected_wins = round(expected_wins,1), wins_over_expected = round(wins_over_expected,1))%>%arrange(desc(expected_wins))

week_4_table<-as.data.frame(week_4_table[,c(2:7)])

asp_ratio <- 1.618 


################### UPDATE PLOT TITLE ##############################
week4_exp_wins<-ggplot(kfl_wins_over_expected,aes(x=reorder(franchise_abbrev,expected_wins),y=expected_wins))+geom_bar(stat="identity")+geom_image(aes(x=franchise_abbrev,y=expected_wins,image=logo), size = 0.05, by = "width", asp = asp_ratio
)+coord_flip()+ggtitle("Expected Wins through Week 10") + xlab("franchise") + ylab("expected wins")+theme(axis.title.x = element_text(size=14),axis.title.y=element_text(size=14),axis.text = element_text(size = 12),plot.title = element_text(size = 18))

week4_exp_wins

################### UPDATE TABLE TITLE ##############################
week_4_table%>%select(franchise_name,logo,expected_wins,wins_over_expected)%>%gt()%>%gt_img_rows(logo) %>%tab_header(title ="Expected Wins through Week 10")%>%
  cols_label(franchise_name="Team Name",logo="",expected_wins="Expected Wins",wins_over_expected="Wins Over Expected")


#theme(aspect.ratio = 1/asp_ratio)

# include aspect ratio in ggsave
#ggsave(
#  "week4_exp_wins.png", week4_exp_wins, 
#  # make the width equivalent to the aspect.ratio
#  height = 20, width = 20 * asp_ratio, dpi = "retina"
#)






```
## Standings through Week 10

Surprisingly the Kraken have dropped to 5-5 and third in the division after a 3-0 start. They really need to view the next 4 weeks as must win if they don't want to be on the outside looking in come playoff time. Both the Supermarket Studs and Tenderloins enjoy winning records at 7-3 and 6-4 to lead the Indiana division.

The Ohio division now has the league best Jeff Slugs at 8-2 followed by the SWAB at 7-3. The bottom of the division has two very competitive spoilers in the Bluecats and Sunday Winners and both teams will look to play spoilers now that their playoff hopes are over. 

Finally, the Glum division is all but wrapped up by Scott's Titans at 7-3 as the only winning team in the division. They will look to defend their title from last year. That is your Week 10 KFL recap. Enjoy your football. 

```{r divison records, echo=FALSE,warning=FALSE,fig.width = 10,fig.height=8,message=FALSE}

kfl_standings_wk4<-kfl_2021_weekly_data%>%filter(week<=week_num)

franchise_id<-1:12
division_name<-c("Ohio","Glum","Ohio","Indiana","Glum","Ohio","Ohio","Indiana","Indiana","Glum","Glum","Indiana")
kfl_division<-as.data.frame(cbind(franchise_id,division_name))
kfl_division$franchise_id<-as.numeric(kfl_division$franchise_id)


kfl_division_data<- sqldf("Select A.*,B.division_name
      from kfl_standings_wk4 A
      left outer join kfl_division B
      on A.Franchise_ID = B.Franchise_ID
      ")

team_df<-kfl_division_data%>%filter(week == 1)%>%select(franchise_id,franchise_name,logo,division_name)

joined_df<-kfl_division_data%>%select(franchise_id,franchise_name,logo,division_name,win)%>%group_by(franchise_name,franchise_id)%>%summarise(
         Wins = length(win[win==1]),
         Losses = length(win[win==0]),
         outcomes = list(win), .groups = "drop") %>% 
     left_join(team_df, by = c("franchise_id" = "franchise_id","franchise_name"="franchise_name")) %>% 
     select(logo,franchise_name, division_name, Wins:outcomes)
 

###################### NEED TO UPDATE THE TITLE HEADER FOR WEEK  ##########################
standings_df<-joined_df%>%select(franchise_name,logo,division_name,Wins,Losses,outcomes)%>%arrange(desc(Wins))%>%gt(groupname_col="division_name")%>%gt_plt_winloss(outcomes,max_wins = 14)%>%gt_img_rows(logo)%>%tab_header(title ="Standings Through Week 10")%>%cols_label(franchise_name="Team",logo="",outcomes="Win-Loss Streak")

standings_df

```

